### 2012-09-04 sections/implementation/execution-workflow.tex
caption consistency with graph legends
penultimate, corrections

A sample JSON query \ac{DNF} term definition is shown in listing
\ref{lst:filter-struct}. In order to be able to make comparisons on the field
offsets of such a term, the comparator needs to know the type of the
comparison and the length of the field offset. This information is parsed by
execution engine once the query is read and is therefore not available at
compile time.  In order to subvert the need to define complex branching
statements, a dedicated comparator is defined for every possible field length
and comparison operation. These comparators are generated using a Python
script. This allows the term definitions to make runtime calls using a
function name derived from the combination of operation type and field length.

@@

\subsubsection{Filter}
\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{filter-fv1-fv2}
    \caption{Filter Stage: NFQL vs NFQL (optimized)}
    \label{fig:fv1-fv2-filter}
  \end{center}
\end{figure}

@@

\begin{figure}[h!]
  \begin{center}
    \includegraphics*[width=0.9\linewidth]{grouper-fv2-bsearch-fv2}
    \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
    \label{fig:fv1-fv2-grouper}
  \end{center}
\end{figure}

>>>>

In order to be able to make comparisons on the field offsets of such a term,
the comparator needs to know the type of the comparison and the length of the
field offset. This information is parsed by execution engine once the query is
read and is therefore not available at compile time.  In order to subvert the
need to define complex branching statements, a dedicated comparator is defined
for every possible field length and comparison operation. A Python script
generates C source code for these comparators at compile time conforming to
the structure shown in listing \ref{filter-struct}. This allows the term
definitions to make runtime calls using a function name derived from the
combination of operation type and field length.

@@

\subsubsection{Filter}
\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{filter-fv1-fv2}
    \caption{Filter Stage: Flowy 2.0 vs NFQL}
    \label{fig:fv1-fv2-filter}
  \end{center}
\end{figure}

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics*[width=0.9\linewidth]{grouper-fv2-bsearch-fv2}
    \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
    \label{fig:fv1-fv2-grouper}
  \end{center}
\end{figure}

### 2012-09-04 sections/implementation/flow-query-intermediate-format.tex
penultimate, corrections

Each stage of the processing pipeline is expressed in the \texttt{JSON} query
as a \ac{DNF} expression. A \ac{DNF} expression is a disjunction of
conjunctive clauses. The elements of the conjuctive clauses are terms. The
clauses in the \ac{DNF} are OR'd together, while the terms in each clause
themselves are AND'd. \texttt{json-c} \cite{jsonc} is used to parse the flow
query file. The pipeline stages of the flow query language are also
encapsulated in Python classes and scripts have been written to facilitate the
front-end parser to serialize these pipeline objects to a \texttt{JSON} flow
query.  The mapping of the query to the structs defined in the execution
engine is not trivial. When reading the \texttt{JSON} query at runtime, the
field names of a NetFlow $v5$ record are read in as strings.  Utility
functions are defined that map the field names to internal struct offsets and
the field types and the operations to internal unique enum members.

@@

The \texttt{JSON} query can also disable the stages at runtime.  This means
that one only has to supply the constructs that one wishes to use.  The
constructs that are not defined in the \texttt{JSON} query are inferred by the
execution engine as a disable request. The execution engine uses disable flags
that are turned on when the query is parsed. These flags are used throughout
the engine to only enable the requested functionality.

>>>>

Each stage of the processing pipeline is expressed in the JSON query as a
\ac{DNF} expression. The elements of the conjuctive clauses are terms.
\texttt{json-c} \cite{jsonc} is used to parse the flow query file. The
pipeline stages of the flow query language are also encapsulated in Python
classes and scripts have been written to facilitate the front-end parser to
serialize these pipeline objects to a JSON flow query.  The mapping
of the query to the structs defined in the execution engine is not trivial.
When reading the JSON query at runtime, the field names of a NetFlow
$v5$ record are read in as strings.  Utility functions are defined that map
the field names to internal struct offsets and the field types and the
operations to internal unique enum members.

@@

The JSON query can also disable the stages at runtime.  This means
that one only has to supply the constructs that one wishes to use.  The
constructs that are not defined in the JSON query are inferred by the
execution engine as a disable request. The execution engine uses disable flags
that are turned on when the query is parsed. These flags are used throughout
the engine to only enable the requested functionality.

### 2012-09-04 sections/introduction.tex
penultimate, corrections

Researchers, service providers and security analysts have long been interested
in network and user behavioral patterns of the traffic crossing the internet
backbone. They want to use this information for the purpose of billing and
mediation, bandwidth provisioning, detecting malicious attacks, network
performance evaluation and overall improvement. Traffic measurement techniques
that have been rapidly evolving in the last decade, have matured enough today
to provide such an insight.

Capturing network flow records has emerged out to be one of the favored
network measurement techniques. In this technique, packets traversing an
obvervation point are not captured raw, instead they are aggregated together
based on some common characteristics. The common characteristics are learnt by
inspecting the packet headers as they cross the observation point.
Flow-records resulting from such an aggregation are then exported to a
collector for further analysis. This reduces the amount of traffic aggregated
at the collector.

NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information
export. NetFlow \cite{rfc3954} is a proprietary network protocol designed by
Cisco Systems. It allows routers to generate and export flow records to a
designated collector. NetFlow v$9$ provides flexibility of user-tailored
export templates, \ac{MPLS} and IPv$6$ support and a larger set of flow keys.
\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF and
deemed to be the logical successor of NetFlow v$9$ on which it is based. The
novelty of the standard lies in its ability to describe record formats at
runtime using templates based on an extensible and well-defined information
model. The data transfer mechanism is also simplistic and extensible by being
unidirectional and transport protocol agnostic.

%The wide applicability of this approach is easily seen from the pervasive use
%of flow records for a vibrant set of network analysis applications. For
%instance, the authors in \cite{mkim:2004} use the flow characteristics in the
%traffic pattern to formalize a detection function that maps traffic patterns
%to different DoS attacks, whereas in \cite{sdominik:2010}, the authors
%use the flow-record data to exploit timing characteristics of webmail clients
%to classify features that could identify webmail traffic from any other
%traffic running over HTTPS.

>>>>

Researchers, service providers and security analysts have long been interested
in network and user behavioral patterns of the traffic crossing the internet
backbone. They want to use this information for the purpose of billing and
mediation, bandwidth provisioning, detecting malicious attacks, network
performance evaluation and overall improvement. Traffic measurement techniques
that have been rapidly evolving in the last decade have today matured enough
to provide such an insight.

Capturing network flow records has emerged out to be one of the favored
network measurement techniques. In this technique, packets traversing an
obvervation point are not captured raw, instead common characteristics of the
packet header are aggregated into flow records. They are then exported to a
collector for further analysis.  This reduces the amount of traffic aggregated
at the collector.

NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information
export. NetFlow \cite{rfc3954} is a network protocol designed by Cisco
Systems, which allows routers to generate and export flow records to a
designated collector. NetFlow v$9$ provides flexibility through user-tailored
export templates, \ac{MPLS} and IPv$6$ support and a larger set of flow keys.
\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF based
on NetFlow v$9$. The novelty of the standard lies in its ability to describe
record formats at runtime using templates based on an extensible and
well-defined information model. The data transfer mechanism is also simplistic
and extensible by being unidirectional and transport protocol agnostic.

### 2012-09-04 sections/relatedwork.tex
penultimate, corrections

%The study of the network traffic patterns began with tools that utilized raw
%packet captures at the monitoring point for traffic measurement.
%\texttt{tcpdump} and \texttt{wireshark} have since then become the de-facto
%tools for packet capture and analysis. \texttt{tcpdump} \cite{tcpdump-manpage}
%is a premier command-line utility that uses the \texttt{libpcap}
%\cite{pcap-manpage} library for packet capture. The power of \texttt{tcpdump}
%comes from the richness of its expressions, the ability to combine them using
%logical connectives and extract specific portions of a packet using filters.
%\texttt{wireshark} \cite{wireshark-manpage} is a GUI application, aimed at
%both journeymen and packet analysis experts. It supports a large number of
%protocols, has a straightforward layout, excellent documentation, and can run
%on all major operating systems.

In recent years, a number of tools have been developed that can capture the
traffic as flow records and use them for network analysis.

Simple analysis of network traffic can be done by a range of graphical
utilities like ntop \cite{ntop:2000}, FlowScan \cite{flowscan:2000}, NfSen
\cite{phaag:2006} and Stager \cite{oslebo:2006}. All these tools understand the
NetFlow format while ntop and Stager can also process \ac{IPFIX}. They use a
round robin database to store flow information, except for Stager which uses a
MySQL database. Graphical presentation is done via a webinterface or in the
case of FlowScan via RRDTool.

\texttt{flow-tools} and \texttt{nfdump} are among the most popular tools used
for analyzing NetFlow data. \texttt{flow-tools} \cite{sromig:2000} is a suite
of programs for capturing and processing NetFlow v5 flow records. It consists
of 24 separate tools that work together by connecting them via UNIX pipes. It
can capture, read, filter, and print flow records internally saved in a
fixed-size format.  \texttt{nfdump} \cite{phaag:2006} is a very similar tool
that uses a different storage format. Flow records are captured using
\texttt{nfcapd} and then processed by \texttt{nfdump}. The power of filtering
rules in both the tools is however mostly limited to absolute comparisons of
flow attributes. As a result, relative comparison amongst different flows or
querying a timing relationship among them is not possible.  SiLK \cite{SiLK} is
a network traffic collection and analysis tool developed and maintained by the
CERT Network Situational Awareness Team (CERT NetSA) at Carnegie Mellon
University.  SiLK is the tool that comes quite close to providing similar
capabilities as provided by \ac{NFQL} and is therefore used as a reference
point to compare the performance of the \ac{NFQL} execution engine in this
paper. The design and implementation of SiLK, however, differs a lot from that
of \ac{NFQL}. For instance, in SiLK there are separate tools to perform the
task of each stage of the \ac{NFQL} processing pipeline. The stage
functionality is not full-fledged though. The grouping and merging operations
can only be performed using an equality operator. This restriction allows the
tool to perform optimization such as using hash tables to perform lookups.
There are also stringent requirements to how flow-data needs to be organized
before it can be piped into a tool. The grouping tool, for instance, assumes
that the to-be supplied input flow data is already sorted on the field column.
These requirements can make it a little cumbersome to design a full-fledged
flow query.  For instance, trying to mimic a \ac{NFQL} query in SiLK sometimes
ends up as a shell script with over a dozen of SiLK tools piped together.

>>>>

In recent years, a number of tools have been developed that can capture the
traffic as flow records and use them for network analysis. Simple analysis of
network traffic can be done by a range of graphical utilities like ntop
\cite{ntop:2000}, FlowScan \cite{flowscan:2000}, NfSen \cite{phaag:2006} and
Stager \cite{oslebo:2006}. All these tools understand the NetFlow format while
ntop and Stager can also process \ac{IPFIX}. They use a round robin database
to store flow information, except for Stager which uses a MySQL database.
Graphical presentation is done via a webinterface or in the case of FlowScan
via RRDTool.

\texttt{flow-tools} and \texttt{nfdump} are among the most popular tools used
for analyzing NetFlow data. \texttt{flow-tools} \cite{sromig:2000} is a suite
of programs for capturing and processing NetFlow v5 flow records. It consists
of 24 separate tools that work together by connecting them via UNIX pipes. It
can capture, read, filter, and print flow records internally saved in a
fixed-size format. \texttt{nfdump} \cite{phaag:2006} is a very similar tool
that uses a different storage format. Flow records are captured using
\texttt{nfcapd} and then processed by \texttt{nfdump}. The power of filtering
rules in both the tools is however mostly limited to absolute comparisons of
flow attributes. As a result, relative comparison amongst different flows or
querying a timing relationship among them is not possible.

SiLK \cite{SiLK} is a network traffic collection and analysis tool developed
and maintained by the CERT Network Situational Awareness Team (CERT NetSA) at
Carnegie Mellon University.  SiLK is the tool that comes quite close to
providing similar capabilities as \ac{NFQL} and is therefore used as a
reference point to compare the performance of the \ac{NFQL} execution engine
in this paper. The design and implementation of SiLK, however, differs a lot
from that of \ac{NFQL}. For instance, in SiLK there are separate tools to
perform the task of each stage of the \ac{NFQL} processing pipeline. The stage
functionality is not full-fledged though. The grouping and merging operations
can only be performed using an equality operator. This restriction allows the
tool to perform optimization such as using hash tables to perform lookups.
\ac{NFQL} on the other hand, provides a much richer set of comparisons
operations, such as \emph{equal, not equal, greater than, less than, greater
or equal, less or equal}. There are also stringent requirements to how
flow-data needs to be organized before it can be piped into a tool. The
grouping tool, for instance, assumes that the to-be supplied input flow data
is already sorted on the field column.  These requirements can make it a
little cumbersome to design a full-fledged flow query.  For instance, trying
to mimic a \ac{NFQL} query in SiLK sometimes ends up as a shell script with
over a dozen of SiLK tools piped together.
